Untangling the link between traits, size and growth rate in plants

### 2015-03-17 ms.tex
update bibtex keys

Functional traits are thought to capture core differences in the strategies
plants use to generate and invest surplus energy
\citep{wright_world_wide_2004, chave_towards_2009, westoby_plant_2002}.
Although most plants have the same basic construction, physiological function
and resource requirements, species differ considerably from one another in
rates of carbon, nitrogen and water uptake, growth, mortality and seed
production. Some of these functional differences have been linked to
underlying variation in relatively simple physiological traits, such as leaf
mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
diameter,  and seed size  \citep{wright_world_wide_2004,chave_towards_2009}.
These traits are easily measured across diverse species, and variation between
species is large compared to variation within species. Data for some traits
now exists for up to 10\% of the world's 250000 plant species
\citep{cornwell_functional_2014}. During the last decade, the idea has taken
hold that the growth strategy and ecology of a plant species might usefully be
characterised simply by knowing its traits.

While the influence of plant traits on elements of plant physiological
function is increasingly understood, attempts at using traits to understand
demographic rates have only met with mixed success
\citep{wright_functional_2010, poorter_are_2008}. In seedlings, the trait LMA
-- a central part of the leaf economics spectrum
\citep{wright_world_wide_2004} --is tightly related to plant relative growth
rate in mass \citep{lambers_inherent_1992, wright_cross_species_2000}. LMA and
its correlate leaf lifespan have also been linked to height growth rate for
small seedlings and saplings \citep{reich_leaf_1992, poorter_leaf_2006}. These
early successes prompted researchers to search for similar relationships in
large plants. However, the results showed that in trees LMA was not correlated
with diameter growth rate  \citep{wright_functional_2010, poorter_are_2008,
herault_functional_2011}. Meanwhile, other traits such as wood density showed
strong relationships to growth in large plants \citep{wright_functional_2010},
but only weak relationships in seedlings  \citep{castro-diez_stem_1998}. Thus
far, most trait growth relationships have been quantified for one or few
traits and at a single life- stage. Yet the picture is emerging is that the
link between traits and growth may be modified by plant size
\citep{ruger_functional_2012}; a finding that challenges the wide-spread
assumption that particular ends of the trait spectrum translate directly into
faster or slower growth.

The challenge of interpreting diverse empirical results linking  -- or not --
traits to growth rate would clearly be easier with clear expectations on what
signal one should expect, given current understandings of how traits influence
plant function. Generating these expectations is a primary, yet perhaps under-
appreciated, role for theory \citep{kokko_modelling_2007}. Current empirical
results suggest the effect of traits on growth changes with size and possibly
also light environment, yet current theory says little about how such
relationships might come about. A widely-used model shows that for seedlings,
mass-based relative growth rate is linearly and negatively related to LMA
\citep{lambers_inherent_1992, cornelissen_seedling_1996,
wright_cross_species_2000}. An extension of the model suggests a similar
relationship may hold at larger sizes \citep{enquist_general_2007}. But as
described above, this prediction is not supported by empirical results.
Meanwhile, theoretical predictions on how other traits should influence growth
are largely absent.

A further problem with existing theory is that the effects of traits are all
realised via influences on mass production\citep{enquist_general_2007},
whereas the effect of traits such as LMA, WD, and Hmax is via the quality of
tissue construction, or the amount of mass allocated to different tissues
\citep{falster_influence_2011}. There are in fact two problems with theory
focussed solely on mass production. The first is that measuring mass
production is only practical for small plants that can be easily harvested.
For larger plants, growth is more often measured either as diameter or height
growth. For example, the CTFS monitors diameter growth in 49 large permanent
plots across forests worldwide \citep{anderson_teixeira_ctfs_forestgeo_2014},
while forestry agencies around the world have likewise recorded data across
large tracts of forest \citep{purves_predictive_2008}. The second problem is
that, as we outline further below, because mass production is only one
component of plant growth; theory focussing solely on mass production will
therefore always be limited.

In this paper we provide a mechanistic framework for describing the potential
effect of traits on growth throughout a plant's life-cycle, linking to
commonly used metrics such as height and diameter growth. To help motivate our
study we re-analyse demographic and trait data from an earlier study
\citep{wright_functional_2010}, to show 1) that the link between traits and
growth is stronger than previously thought, 2) that the influence of three
prominent traits on growth rate is moderated by plant size, and 3) that the
impact of size on the nature of relationships varies among traits (see also
\citet{ruger_functional_2012}. We then show how these results align with new
set of theoretical predictions. Moreover, we show how traits moderate plant
responses to light environment and also determine shade tolerance, supporting
diverse empirical results \citep{ruger_functional_2012, poorter_leaf_2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, we hope to provide a solid theoretical foundation for trait
ecology and thus provide a platform for understanding growth across diverse
species around the world.

@@

To help illustrate the empirical patterns, we begin with a fresh analysis of
growth dynamics in the long-term plot at Barro Colorado Island in Panama,
where the growth rates for over 100 species have been recorded over the last
30 years. Data from a long term plot \citep{condit_barro_2012} are combined
with results of experiment on seedlings \citep{kitajima_leaf_2013}, to better understand
size-related patterns of growth. A prior analysis by
\citet{wright_functional_2010} found a moderate effect of wood density on
growth rate, but almost no effect of LMA, and only a weak effect of HMAX. More recently,
\citet{ruger_functional_2012} showed

>>>>

Functional traits are thought to capture core differences in the strategies
plants use to generate and invest surplus energy
\citep{wright_world-2004, chave-2009, westoby-2002}.
Although most plants have the same basic construction, physiological function
and resource requirements, species differ considerably from one another in
rates of carbon, nitrogen and water uptake, growth, mortality and seed
production. Some of these functional differences have been linked to
underlying variation in relatively simple physiological traits, such as leaf
mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
diameter,  and seed size  \citep{wright_world-2004,chave-2009}.
These traits are easily measured across diverse species, and variation between
species is large compared to variation within species. Data for some traits
now exists for up to 10\% of the world's 250000 plant species
\citep{cornwell-2014}. During the last decade, the idea has taken
hold that the growth strategy and ecology of a plant species might usefully be
characterised simply by knowing its traits.

While the influence of plant traits on elements of plant physiological
function is increasingly understood, attempts at using traits to understand
demographic rates have only met with mixed success
\citep{wright-2010, poorter-2008}. In seedlings, the trait LMA
-- a central part of the leaf economics spectrum
\citep{wright_world-2004} --is tightly related to plant relative growth
rate in mass \citep{lambers-1992, wright_cross-2000}. LMA and
its correlate leaf lifespan have also been linked to height growth rate for
small seedlings and saplings \citep{reich-1992, poorter-2006}. These
early successes prompted researchers to search for similar relationships in
large plants. However, the results showed that in trees LMA was not correlated
with diameter growth rate  \citep{wright-2010, poorter-2008,
herault-2011}. Meanwhile, other traits such as wood density showed
strong relationships to growth in large plants \citep{wright-2010},
but only weak relationships in seedlings  \citep{castro-diez-1998}. Thus
far, most trait growth relationships have been quantified for one or few
traits and at a single life- stage. Yet the picture is emerging is that the
link between traits and growth may be modified by plant size
\citep{ruger-2012}; a finding that challenges the wide-spread
assumption that particular ends of the trait spectrum translate directly into
faster or slower growth.

The challenge of interpreting diverse empirical results linking  -- or not --
traits to growth rate would clearly be easier with clear expectations on what
signal one should expect, given current understandings of how traits influence
plant function. Generating these expectations is a primary, yet perhaps under-
appreciated, role for theory \citep{kokko-2007}. Current empirical
results suggest the effect of traits on growth changes with size and possibly
also light environment, yet current theory says little about how such
relationships might come about. A widely-used model shows that for seedlings,
mass-based relative growth rate is linearly and negatively related to LMA
\citep{lambers-1992, cornelissen-1996,
wright_cross-2000}. An extension of the model suggests a similar
relationship may hold at larger sizes \citep{enquist-2007}. But as
described above, this prediction is not supported by empirical results.
Meanwhile, theoretical predictions on how other traits should influence growth
are largely absent.

A further problem with existing theory is that the effects of traits are all
realised via influences on mass production\citep{enquist-2007},
whereas the effect of traits such as LMA, WD, and Hmax is via the quality of
tissue construction, or the amount of mass allocated to different tissues
\citep{falster-2011}. There are in fact two problems with theory
focussed solely on mass production. The first is that measuring mass
production is only practical for small plants that can be easily harvested.
For larger plants, growth is more often measured either as diameter or height
growth. For example, the CTFS monitors diameter growth in 49 large permanent
plots across forests worldwide \citep{anderson_teixeira_ctfs-2014},
while forestry agencies around the world have likewise recorded data across
large tracts of forest \citep{purves-2008}. The second problem is
that, as we outline further below, because mass production is only one
component of plant growth; theory focussing solely on mass production will
therefore always be limited.

In this paper we provide a mechanistic framework for describing the potential
effect of traits on growth throughout a plant's life-cycle, linking to
commonly used metrics such as height and diameter growth. To help motivate our
study we re-analyse demographic and trait data from an earlier study
\citep{wright-2010}, to show 1) that the link between traits and
growth is stronger than previously thought, 2) that the influence of three
prominent traits on growth rate is moderated by plant size, and 3) that the
impact of size on the nature of relationships varies among traits (see also
\citet{ruger-2012}. We then show how these results align with new
set of theoretical predictions. Moreover, we show how traits moderate plant
responses to light environment and also determine shade tolerance, supporting
diverse empirical results \citep{ruger-2012, poorter-2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, we hope to provide a solid theoretical foundation for trait
ecology and thus provide a platform for understanding growth across diverse
species around the world.

@@

To help illustrate the empirical patterns, we begin with a fresh analysis of
growth dynamics in the long-term plot at Barro Colorado Island in Panama,
where the growth rates for over 100 species have been recorded over the last
30 years. Data from a long term plot \citep{condit-2012} are combined
with results of experiment on seedlings \citep{kitajima-2013}, to better understand
size-related patterns of growth. A prior analysis by
\citet{wright-2010} found a moderate effect of wood density on
growth rate, but almost no effect of LMA, and only a weak effect of HMAX. More recently,
\citet{ruger-2012} showed

### 2015-03-17 notes/ms_short.tex
update bibtex keys

Plant growth is the foundation of the terrestrial biosphere, powering
food-webs, the carbon cycle and much of human
activity \citep{raupach_carbon_2010}. Plant species differ in many
functional traits, including aspects of tissue construction and energy
allocation, that drive differences in mechanical strength, biomass
partitioning, and rates of photosynthesis and tissue
turnover\citep{wright_world_wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
such traits influence whole-plant growth, with the simple relationships
predicted by existing theory often lacking empirical
support\citep{wright_functional_2010,poorter_are_2008}. Here we
provide a new set of theoretical predictions, validated using 20 years
of growth data\citep{condit_barro_2012,kitajima_leaf_2013}, showing
that the influence of three prominent traits on growth rate is moderated
by plant size, and that the impact of size on the nature of
relationships varies among traits. We extend a widely-used theoretical
model that links growth rate in seedlings with a single leaf
trait\citep{lambers_inherent_1992,wright_cross_species_2000} to
explicitly include influences of size, light environment, and three
other prominent traits. We find the influence of leaf construction cost
(\(\phi\)) on growth is strong in small plants but weakens with size. In
contrast, the influence of stem construction cost (\(\rho\)) peaks at
intermediate sizes, while the influence of height at maturation
(\(h_m\)) strengthens with size. Our theoretical model suggests the
change in trait-growth relationships with size arises due to changes in
biomass partitioning and reproductive allocation. Moreover, we show how
traits moderate plant responses to light environment and also determine
shade tolerance, supporting diverse empirical
results\citep{ruger_functional_2012,poorter_leaf_2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, our results provide a solid theoretical foundation for
trait ecology and thus provide a platform for understanding growth
across diverse species around the world.

@@

Functional traits capture core differences in the strategies plants use
to generate and invest surplus energy\citep{wright_world_wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
traits now exists for many of the world's 250000 plant
species\citep{cornwell_functional_2014}; yet the effect of these traits
on growth remains poorly understood. Prominent traits such as \(\phi\)
(also known as leaf mass per area) are widely considered as indicators
of growth rate\citep{reich_world_wide_2014}, because cheaper leaf
construction should enable faster canopy
expansion\citep{blackman_compound_1919,westoby_time_2000}. Indeed,
growth rate in seedlings is strongly and negatively correlated with
\(\phi\)\citep{cornelissen_seedling_1996,wright_cross_species_2000}
(Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
might also give a growth advantage in larger
plants\citep{enquist_general_2007}, an increasing number of empirical
studies have failed to detect any such
effect\citep{wright_functional_2010, @herault_functional_2011,
@poorter_are_2008}. Meanwhile, expectations on how other prominent
traits might influence growth are lacking. The growing discord between
theoretical expectation and empirical result thus indicates a strong
need for a new theoretical framework through which the influence of
traits and size on plant growth can be understood.

Here we extend a widely-used theoretical model that links growth rate in
seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross_species_2000}, to explicitly include influences of
size, light environment, and other prominent traits. This extension
allows us to generate novel predictions on the link between four key
traits and plant demography (Table \ref{tab:predictions}). Predictions
are compared against data from both a long-term
plot\citep{condit_barro_2012} and field
experiment\citep{kitajima_leaf_2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{S-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies\citep{wright_functional_2010}. The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
\ref{tab:predictions}) showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

We begin with a standard model for the amount of biomass available for
growth, \(dP / \textrm{d}t\), given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant\citep{falster_influence_2011}:

@@

Current theory emphasises the relationship between \(\phi\) and relative
growth in mass\citep{enquist_general_2007},
\(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area (\(A_\textrm{net} = Y(A - r_\textrm{l})\)), making relative growth
rate a linear function of \(\phi\):

@@

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to \(\phi\)\citep{wright_cross_species_2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height (\(h\)) and stem cross-sectional area (\(a_\textrm{st}\)) or
diameter \(D\).

@@

Mathematically, height growth can be decomposed into a product of
physiologically relevant terms\citep{falster_influence_2011}:

@@

The first term on the right of eq \ref{eq:dhdt},
\(\textrm{d}h / \textrm{d}a_\textrm{l}\), is the growth in plant height
per unit growth in total leaf area; accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension\citep{poorter_architecture_2006}.

@@

To address the first of these challenges, we use a functional balance
model\citep{yokozawa_foliage_1995,falster_influence_2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{f-assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

This new growth model captures the intrinsically size-dependent nature
of plant growth (Table \ref{tab:predictions}a), recovering well-known
empirical patterns\citep{sillett_increasing_2010, king_size-related_2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size\citep{sillett_increasing_2010,
@stephenson_rate_2014}, due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve\citep{herault_functional_2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates\citep{iida_linking_2014}.

@@

By definition, the main effect of increasing \(\phi\) is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of \(\phi\) on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction\citep{wright_world_wide_2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dPdt} to \(\phi\). We use an allometric scaling relationship of
the form \(k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}\), which has been
shown to describe patterns across diverse
ecosystems\citep{wright_world_wide_2004} (Fig. \ref{fS-leaf}).

@@

As for leaf, cheaper stem construction (lower \(\rho\)) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
\(\rho\) are less understood than for \(\phi\). One possibility is that
lower \(\rho\) decreases mechanical strength resulting in higher
mortality\citep{chave_towards_2009,wright_functional_2010}. Under
this scenario, decreasing \(\rho\) will always provide a growth
advantage, because the costs of low \(\rho\) are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower \(\rho\) may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with \(\rho\),
i.e.\(k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}\), with \(\beta_5 > 1\).
Under this scenario, decreasing \(\rho\) provides an advantage in height
growth only when the benefits of low \(\rho\) outweigh the costs. This
effect is strongest at intermediate sizes. The decline in height growth
at low \(\rho\) with size is more marked than for either stem basal-area
and diameter growth, because high sapwood turnover also contributes to
this growth measure. Thus diameter growth continues, even as net
production approaches zero.

@@

Two other important predictions arise directly from linking traits to
growth rate (Table \ref{tab:predictions}c). First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low \(rho\) increase growth more substantially with incraeses in
light\citep{ruger_functional_2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low \(\phi\) are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
\(rho\) actually decreases as size
increases\citep{ruger_functional_2012}. In contrast, growth rate tends
to increase in size for the species with highest
\(\rho\)\citep{ruger_functional_2012}.

Variation in \(\phi\) also moderates the response of growth to changes
in light, with species having the lowest \(rho\) being most responsive.
However, unlike for \(\rho\), the effect appears only for the smallest
size classes. Accordingly, empirical analyses have detected such a link
in small plants but not for plants \textgreater{} 0.01m
dbh\citep{ruger_functional_2012}.

Second, species with low \(\phi\) and low \(\rho\) are predicted to be
less shade tolerant (Fig. \ref{f-wplcp}). At low \(\phi\) (\(\rho\)),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
\citep{givnish_adaptation_1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both \(\phi\) and
\(\rho\) have been linked to shade tolerance\citep{poorter_leaf_2006, lusk_ontogenetic_2008,osunkoya_light_1996}.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  traits do effect growth rate, but not in the simple linear fashion
  previously anticipated.
\item
  Relationships change with size due to changes in biomass partitioning
  and reproductive allocation.
\item
  explains why LMA does not show relationship to growth among larger
  plants. Does not mean unimportant for growth, just effects
  concentrated at small sizes.
\item
  in fact, explains plasticity in LMA with size, where LMA tends to
  increase with size \citep{thomas_asymptotic_1999;
  @thomas_photosynthetic_2010}
\item
  Framework can extend to other traits (:

@@

Note, scaling relationship reversed from \citep{falster_influence_2011}.

@@

We follow the model of\citep{yokozawa_foliage_1995} describing the
vertical distribution of leaf area within the crowns of individual
plants. This model can account for a variety of canopy profiles through
a single parameter \(\eta\). Setting \(\eta=1\) results in a conical
canopy, as seen in many conifers, while higher values, e.g. \(\eta=12\)
, give a top-weighted canopy profile similar to those seen among
angiosperms. Let \(S(z,h)\) be the sapwood area at height \(z\) for a
plant with top height \(h\). Following Yokozawa and Hara (1995) we
assume a relationship between \(S(z,h)\) and height such that

@@

We also assume that each unit of sapwood area supports a fixed area of
leaf (the pipe model, \citep{shinozaki_quantitative_1964}), so that the total
canopy area of a plant relates to basal sapwood area \(S(0,h)\):

@@

where
\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa_foliage_1995}.
Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
expression for sapwood mass as a function leaf area and height:

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_tradeoff.pdf}
\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
Data from \citep{wright_world_wide_2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship.\label{fS-leaf}}
\end{figure}

>>>>

Plant growth is the foundation of the terrestrial biosphere, powering
food-webs, the carbon cycle and much of human
activity \citep{raupach-2010}. Plant species differ in many
functional traits, including aspects of tissue construction and energy
allocation, that drive differences in mechanical strength, biomass
partitioning, and rates of photosynthesis and tissue
turnover\citep{wright_world-2004,chave-2009, westoby-2002}. Yet, it remains unclear how -- and even if --
such traits influence whole-plant growth, with the simple relationships
predicted by existing theory often lacking empirical
support\citep{wright-2010,poorter-2008}. Here we
provide a new set of theoretical predictions, validated using 20 years
of growth data\citep{condit-2012,kitajima-2013}, showing
that the influence of three prominent traits on growth rate is moderated
by plant size, and that the impact of size on the nature of
relationships varies among traits. We extend a widely-used theoretical
model that links growth rate in seedlings with a single leaf
trait\citep{lambers-1992,wright_cross-2000} to
explicitly include influences of size, light environment, and three
other prominent traits. We find the influence of leaf construction cost
(\(\phi\)) on growth is strong in small plants but weakens with size. In
contrast, the influence of stem construction cost (\(\rho\)) peaks at
intermediate sizes, while the influence of height at maturation
(\(h_m\)) strengthens with size. Our theoretical model suggests the
change in trait-growth relationships with size arises due to changes in
biomass partitioning and reproductive allocation. Moreover, we show how
traits moderate plant responses to light environment and also determine
shade tolerance, supporting diverse empirical
results\citep{ruger-2012,poorter-2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, our results provide a solid theoretical foundation for
trait ecology and thus provide a platform for understanding growth
across diverse species around the world.

@@

Functional traits capture core differences in the strategies plants use
to generate and invest surplus energy\citep{wright_world-2004, chave-2009,westoby-2002}. Data for prominent
traits now exists for many of the world's 250000 plant
species\citep{cornwell-2014}; yet the effect of these traits
on growth remains poorly understood. Prominent traits such as \(\phi\)
(also known as leaf mass per area) are widely considered as indicators
of growth rate\citep{reich_world-2014}, because cheaper leaf
construction should enable faster canopy
expansion\citep{blackman-1919,westoby-2000}. Indeed,
growth rate in seedlings is strongly and negatively correlated with
\(\phi\)\citep{cornelissen-1996,wright_cross-2000}
(Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
might also give a growth advantage in larger
plants\citep{enquist-2007}, an increasing number of empirical
studies have failed to detect any such
effect\citep{wright-2010, @herault-2011,
@poorter-2008}. Meanwhile, expectations on how other prominent
traits might influence growth are lacking. The growing discord between
theoretical expectation and empirical result thus indicates a strong
need for a new theoretical framework through which the influence of
traits and size on plant growth can be understood.

Here we extend a widely-used theoretical model that links growth rate in
seedlings to \(\phi\)\citep{lambers-1992, wright_cross-2000}, to explicitly include influences of
size, light environment, and other prominent traits. This extension
allows us to generate novel predictions on the link between four key
traits and plant demography (Table \ref{tab:predictions}). Predictions
are compared against data from both a long-term
plot\citep{condit-2012} and field
experiment\citep{kitajima-2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{S-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies\citep{wright-2010}. The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
\ref{tab:predictions}) showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

We begin with a standard model for the amount of biomass available for
growth, \(dP / \textrm{d}t\), given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant\citep{falster-2011}:

@@

Current theory emphasises the relationship between \(\phi\) and relative
growth in mass\citep{enquist-2007},
\(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area (\(A_\textrm{net} = Y(A - r_\textrm{l})\)), making relative growth
rate a linear function of \(\phi\):

@@

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to \(\phi\)\citep{wright_cross-2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height (\(h\)) and stem cross-sectional area (\(a_\textrm{st}\)) or
diameter \(D\).

@@

Mathematically, height growth can be decomposed into a product of
physiologically relevant terms\citep{falster-2011}:

@@

The first term on the right of eq \ref{eq:dhdt},
\(\textrm{d}h / \textrm{d}a_\textrm{l}\), is the growth in plant height
per unit growth in total leaf area; accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension\citep{poorter-2006}.

@@

To address the first of these challenges, we use a functional balance
model\citep{yokozawa-1995,falster-2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{f-assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

This new growth model captures the intrinsically size-dependent nature
of plant growth (Table \ref{tab:predictions}a), recovering well-known
empirical patterns\citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size\citep{sillett-2010,
@stephenson-2014}, due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve\citep{herault-2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates\citep{iida-2014}.

@@

By definition, the main effect of increasing \(\phi\) is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of \(\phi\) on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction\citep{wright_world-2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dPdt} to \(\phi\). We use an allometric scaling relationship of
the form \(k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}\), which has been
shown to describe patterns across diverse
ecosystems\citep{wright_world-2004} (Fig. \ref{fS-leaf}).

@@

As for leaf, cheaper stem construction (lower \(\rho\)) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
\(\rho\) are less understood than for \(\phi\). One possibility is that
lower \(\rho\) decreases mechanical strength resulting in higher
mortality\citep{chave-2009,wright-2010}. Under
this scenario, decreasing \(\rho\) will always provide a growth
advantage, because the costs of low \(\rho\) are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower \(\rho\) may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with \(\rho\),
i.e.\(k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}\), with \(\beta_5 > 1\).
Under this scenario, decreasing \(\rho\) provides an advantage in height
growth only when the benefits of low \(\rho\) outweigh the costs. This
effect is strongest at intermediate sizes. The decline in height growth
at low \(\rho\) with size is more marked than for either stem basal-area
and diameter growth, because high sapwood turnover also contributes to
this growth measure. Thus diameter growth continues, even as net
production approaches zero.

@@

Two other important predictions arise directly from linking traits to
growth rate (Table \ref{tab:predictions}c). First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low \(rho\) increase growth more substantially with incraeses in
light\citep{ruger-2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low \(\phi\) are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
\(rho\) actually decreases as size
increases\citep{ruger-2012}. In contrast, growth rate tends
to increase in size for the species with highest
\(\rho\)\citep{ruger-2012}.

Variation in \(\phi\) also moderates the response of growth to changes
in light, with species having the lowest \(rho\) being most responsive.
However, unlike for \(\rho\), the effect appears only for the smallest
size classes. Accordingly, empirical analyses have detected such a link
in small plants but not for plants \textgreater{} 0.01m
dbh\citep{ruger-2012}.

Second, species with low \(\phi\) and low \(\rho\) are predicted to be
less shade tolerant (Fig. \ref{f-wplcp}). At low \(\phi\) (\(\rho\)),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
\citep{givnish-1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both \(\phi\) and
\(\rho\) have been linked to shade tolerance\citep{poorter-2006, lusk-2008,osunkoya-1996}.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  traits do effect growth rate, but not in the simple linear fashion
  previously anticipated.
\item
  Relationships change with size due to changes in biomass partitioning
  and reproductive allocation.
\item
  explains why LMA does not show relationship to growth among larger
  plants. Does not mean unimportant for growth, just effects
  concentrated at small sizes.
\item
  in fact, explains plasticity in LMA with size, where LMA tends to
  increase with size \citep{thomas-1999;
  @thomas-2010}
\item
  Framework can extend to other traits (:

@@

Note, scaling relationship reversed from \citep{falster-2011}.

@@

We follow the model of\citep{yokozawa-1995} describing the
vertical distribution of leaf area within the crowns of individual
plants. This model can account for a variety of canopy profiles through
a single parameter \(\eta\). Setting \(\eta=1\) results in a conical
canopy, as seen in many conifers, while higher values, e.g. \(\eta=12\)
, give a top-weighted canopy profile similar to those seen among
angiosperms. Let \(S(z,h)\) be the sapwood area at height \(z\) for a
plant with top height \(h\). Following Yokozawa and Hara (1995) we
assume a relationship between \(S(z,h)\) and height such that

@@

We also assume that each unit of sapwood area supports a fixed area of
leaf (the pipe model, \citep{shinozaki-1964}), so that the total
canopy area of a plant relates to basal sapwood area \(S(0,h)\):

@@

where
\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa-1995}.
Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
expression for sapwood mass as a function leaf area and height:

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_tradeoff.pdf}
\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
Data from \citep{wright_world-2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship.\label{fS-leaf}}
\end{figure}
